---
title: "Module 04 Lab"
author: "Daniel Jackson"
date: "September 19th, 2023"
output: pdf_document
---

# Lab Data
```{r lab data, include=TRUE}
download.file("http://www.openintro.org/stat/data/bdims.RData", destfile = "bdims.RData")
load("bdims.RData")

head(bdims)

mdims <- subset(bdims, sex == 1)
fdims <- subset(bdims, sex == 0)
```

# Exercise 1
```{r Exercise 1}
library(ggplot2)
# Histogram for males
ggplot(mdims,
       mapping = aes(x = hgt)) +
  geom_histogram(color = 'white', fill = 'red') +
  labs(x = 'Height in cm',
       y = 'Count',
       title = 'Counts of Males by Height in cm')

# Histograms for females
ggplot(fdims,
       mapping = aes(x = hgt)) +
  geom_histogram(color = 'white', fill = 'blue') +
  labs(x = 'Height in cm',
       y = 'Count',
       title = 'Counts of Females by Height in cm')

# One histogram with height by males and females
ggplot(bdims,
       mapping = aes(x = hgt, fill = sex)) +
  geom_histogram(color = 'white') +
  labs(x = 'Height in cm',
       y = 'Count',
       title = 'Counts of Males vs. Females by Height in cm')
```

 Both plots look relatively normal. Both distributions are unimodial and the males histogram looks almost symmetric. Both have a bell-shape to them.

# Exercise 2
```{r Exercise 2}
# Working with woman's height, find mean and sd:
fhgtmean <- mean(fdims$hgt)
fhgtsd   <- sd(fdims$hgt)

# Make density histogram to use as backdrop and overlay a normal probability curve
hist(fdims$hgt, probability = TRUE, ylim = c(0, 0.06))
x <- 140:190
y <- dnorm(x = x, mean = fhgtmean, sd = fhgtsd)
lines(x = x, y = y, col = "blue")
```

Based on this density histogram, I would say the data follows a nearly normal distribution.

# Exercise 3
```{r Exercise 3}
# We can also construct a normal probability plot to check if distribution is nearly normal. This plot is also called a Q-Q plot for , "quantile-quantile". A data set that is nearly normal will result in a probability plot where the points closely 
## follow the line.

qqnorm(fdims$hgt)
qqline(fdims$hgt)

# What do probability plots look like for data that I know came from a normal distribution? We can answer this by simulating data from a normal distribution using rnorm.

sim_norm <- rnorm(n = length(fdims$hgt), mean = fhgtmean, sd = fhgtsd)
qqnorm(sim_norm)
qqline(sim_norm)
```

Not all points fall on the line. But it seems that more points are closer to the line compared to the probability plot for the real data.The points near the line seem to be more linear in the probability plot for sim_norm compared to the real data.

# Exercise 4
Even better than comparing the original plot to a single plot generated from a normal distribution is to compare it to many more plots using the following function.
```{r Exercise 4}
qqnormsim(fdims$hgt)
```

All of the outputs look relatively linear, and very similar to real data.

# Exercise 5
Using the same technique, determine whether or not female weights appear to come from a normal distribution.
```{r Exercise 5}
ggplot(fdims,
       mapping = aes(x = wgt)) +
  geom_histogram(color = 'white', fill = 'blue') +
  labs(x = 'Weight in kg',
       y = 'Count',
       title = 'Counts of Females by Weight in kg')

# Working with woman's weight, find mean and sd:
fwgtmean <- mean(fdims$wgt)
fwgtsd   <- sd(fdims$wgt)


# Make density histogram to use as backdrop and overlay a normal probability curve
min(fdims$wgt)
max(fdims$wgt)

hist(fdims$wgt, probability = TRUE, ylim = c(0, 0.06))
x <- 42:105.2
y <- dnorm(x = x, mean = fwgtmean, sd = fwgtsd)
lines(x = x, y = y, col = "blue")

# Based on the distribution from the density histogram, it does not look like the 
# distribution is normal. Looks more right-skewed than symmetric.

# Let's look at normal probability plot:

qqnorm(fdims$wgt)
qqline(fdims$wgt)

# The probablity plot is concave up which further confirms my thoughts that the 
# distribution is right-skewed.

# Let's simulate data from a normal distribution using rnorm.
sim_norm_wgt <- rnorm(n = length(fdims$wgt), mean = fwgtmean, sd = fwgtsd)
qqnorm(sim_norm_wgt)
qqline(sim_norm_wgt)
```

This data looks s-shaped more than linear. Therefore, I am going to say that the female weight distribution is right-skewed and not normal.

# Exercise 6
What is the probability that a randomly chosen young adult female is taller than 6 feet (about 182 cm)?
```{r Exercise 6}
print(1 - pnorm(q = 182, mean = fhgtmean, sd = fhgtsd))
# Returned 0.0044 which is 0.44%

# Assuming a normal distribution has allowed us to calculate a theoretical probability.If we want to calculate the probability empirically, we simply need to determine how many observations fall above 182 then divide this number by the total sample size.

print(sum(fdims$hgt > 182) / length(fdims$hgt))
# This returned 0.0038 which is 0.38%
```
Probability that a randomly chose young adult femaile is taller than 6 feet is 0.44%

Question 1: What is the probability that a randomly chosen young adult female is smaller than 5 feet 7 inches (about 170 cm)?
```{r Exercise 6 Question 1}
print(pnorm(170, mean = fhgtmean, sd = fhgtsd))
# Returned 0.7833 which is 78.33%

# Calculate probability empirically
print(sum(fdims$hgt < 170) / length(fdims$hgt))
# Returned 0.7538 which is 75.38%

# Difference between two probabilities:
print(78.33 - 75.38)
# Returned 2.95
```
Probability is 78.33%.
Probability empirically is 75.38%
Difference between the two probabilities is 2.95%

Question 2: What is the probability that a randomly chosen young adult female weighs more than 55 kg? 
```{r Exercise 6 Question 2}
print(1 - (pnorm(55, mean = fwgtmean, sd = fwgtsd)))
# This returned 0.7199 which is 71.99%

# Calculate probability empirically
print(sum(fdims$wgt > 55) / length(fdims$wgt))
# This returned 0.6923 which is 69.23%

# Difference between two probabilities:
print(71.99 - 69.23)
# Returned 2.76
```
Probability is 71.99%.
Probability empirically is 69.23%
Difference between the two probabilities is 2.76%

Weight had a closer agreement between two methods. However, height is more normal than weight, so I believe the probabilities of height are more accurate.

# On Your Own Lab
1.)
a.) The histogram for female biiliac (pelvic) diameter (bii.di) belongs to normal probability plot letter ____.
```{r On your own lab 1a}
qqnorm(fdims$bii.di)
qqline(fdims$bii.di)
```

Belongs to plot B

b.) The histogram for female elbow diameter (elb.di) belongs to normal probability plot letter ____.
```{r On your own lab 1b}
qqnorm(fdims$elb.di)
qqline(fdims$elb.di)
```

Belongs to plot C

c.) The histogram for general age (age) belongs to normal probability plot letter ____.
```{r On your own lab 1c}
qqnorm(fdims$age)
qqline(fdims$age)
```

Belongs to plot D

d.) The histogram for female chest depth (che.de) belongs to normal probability plot letter ____.
```{r On your own lab 1d}
qqnorm(fdims$che.de)
qqline(fdims$che.de)
```

Belongs to plot A

2.) This occurs because these distributions are not normal. The more ouutliers there are, the more skewed the data.

3.)
```{r On your own lab 3}
qqnorm(fdims$kne.di)
qqline(fdims$kne.di)
# Based on probability plot being concave up, this looks right-skewed.

# Create histogram to confirm conclusion.
ggplot(fdims,
       mapping = aes(x = kne.di)) +
  geom_histogram(color = 'white', fill = 'blue') +
  labs(x = 'Knee Diameter',
       y = 'Count',
       title = "Counts of Female's Knee Diameter")
```

Confirmed with histogram that data is right-skewed.