---
title: "Module 04 Exercises"
author: "Daniel Jackson"
date: "September 19th, 2023"
output: pdf_document
---

Note: I could not convert plots properly using R Markdown. There will be a word document submitted with code and plot for Exercise 4.2 a, b, c, and d.

# Exercise 4.2
a.) Z > -1.13
```{r Exercise 4.2 a}
# Code to get answer:
print(1 - pnorm(-1.13, mean = 0, sd = 1))
```
Answer: 87.1%

b.) Z < 0.18
```{r Exercise 4.2 b}
# Code to get answer:
print(pnorm(0.18, mean = 0, sd = 1))
```
Answer: 57.1%

c.) Z > 8
```{r Exercise 4.2 c1}
# Code to get answer:
print(1 - pnorm(8, mean = 0, sd = 1))
```
Answer: approximately 0%

d.) |Z| < 0.5
```{r Exercise 4.2 d1}
# Code to get answer:
1 - (1 - pnorm(0.5)) - pnorm(-0.5)
```
Answer is 38.3%

# Exercise 4.4
a.) 
Men, Ages 30-34 N($\mu$ = 4313, $\sigma$ = 583)

Women, Ages 25-29 N($\mu$ = 5261, $\sigma$ = 807)

b.)

```{r Exercise 4.4 b1}
# Leo's Z-score
leo_z = (4948 - 4313) / 583
print(leo_z)

# Mary's Z-score
mary_z = (5513 - 5261) / 807
print(mary_z)
```
Leo is 1.09 standard deviations from mean time of 4313 in his group.
Mary is 0.3123 standard deviations from mean time of 5261 in her group.

c.)

```{r Exercise 4.4 c1}
# Leo:
print(pnorm(1.09))
```
Leo finished slower than over 86% of finishers in his group.

```{r Exercise 4.4 c2}
# Mary:
print(pnorm(0.3123))
```
Mary finished slower than over 62% of finishers in her group.

Mary ranked better in her respective group compared to Leo, as she was less slower in her group compared to Leo.


d.)
```{r Exercise 4.4 d}
print(1 - pnorm(1.09))
```
Leo finished faster than over 13% of finishers.

e.)
```{r Exercise 4.4 e}

print(1 - pnorm(0.3123))
```
Mary finished faster than over 37% of finishers.

f.) 
If the distribution was not normal, our answers from part b to part e would absolutely change. All of the calculations that were doing were under the requirements of the distribution being normal. 

# Exercise 4.8
a.)
Z-score < 0:
```{r Exercise 4.8 a}
zero_z = pnorm(0, mean = 14.7, sd = 33)
print(zero_z)
```
This portfolio loses money approximately 32.8% of years

b.)
Cutoff for highest 15% of annual returns.
100% - 15% = 85th percentile
```{r Exercise 4.8 b}
percentage = qnorm(0.85, mean = 14.7, sd = 33)
print(percentage)
```
The cutoff for highest 15% of annual returns is 48.9%.